Data exploration, quality control and statistical analysis of ChIP-exo experiments

### 2016-04-12 rmarkdown/paper.Rnw
preliminary version of the paper, pre -change of sections

    ChIP-exo is a modification of the ChIP-Seq protocol for high
    resolution mapping of transcription factor binding sites. Although
    many aspect of the ChIP-exo data analysis are similar to those of
    ChIP-Seq, ChIP-exo presents a number of unique challenges. We
    present a quality control pipeline that analyzes a ChIP-exo
    experiment's strand imbalance, enrichment and library
    complexity. Assessment of these characteristics are facilitated
    through diagnostic plots and summary statistics calculated over
    regions of the genome with varying levels of coverage.

@@

\begin{keyword}
  \kwd{ChIP-exo}
  \kwd{Quality Control}
  \kwd{ChIP-Seq}
  \kwd{Spatial Resolution}
  \kwd{Transcription Factor}  
  \kwd{Binding Site Identification on High-Res}
  \kwd{Deconvolution}
\end{keyword}

@@

@@

% Reffers to... First, DNA libraries... I think that these might be
% too much details for the introduction. How about discussing them in
% the results when we investigate the related

@@

\begin{table}[h!]
%% need to fix format
  \centering
<<table1,include = FALSE,echo = FALSE,eval = TRUE>>=
  library(data.table)
  load("../data/paper_tables/Landick_rif_summary.RData")
  load("../data/paper_tables/Landick_aero_summary.RData")
  load("../data/paper_tables/pugh_ctcf.RData")
  load("../data/paper_tables/Carroll_mm9_FoxA1.RData")
  load("../data/paper_tables/Meijsing_hg19_GR.RData")
  load("../data/paper_tables/Carroll_hg19_ER.RData")
  load("../data/paper_tables/Venters_hg19_TBP.RData")
  landick_aero[,Organism := "E.Coli"]
  landick_aero[,Replicate := c(1,2,1,2)]
  landick_aero[,TF := "Sig70"]
  landick_aero[ ,cond := c(rep("Aerobic",2),rep("Anaerobic", 2))]
  landick_rif[,Organism := "E.Coli"]
  landick_rif[,Replicate := c(1,1,2,2)]
  landick_rif[,TF := "Sig70"]
  landick_rif[ ,cond := rep(c("Rif-0min","Rif-20min"),each = 2)]
  pugh_ctcf[,Organism := "Human"]
  pugh_ctcf[,Replicate := 1]
  pugh_ctcf[,TF := "CTCF"]
  pugh_ctcf[ ,cond := "HeLA"]
  carroll_mm9_FoxA1[,Organism := "Mouse"]
  carroll_mm9_FoxA1[,Replicate := c(3,1,2)]
  carroll_mm9_FoxA1[,TF := "Fox A1"]
  carroll_mm9_FoxA1[ ,cond := "Mouse Liver"]
  meijing_hg19_GR[,Organism := "Human"]
  meijing_hg19_GR[,Replicate := 1]
  meijing_hg19_GR[, TF := "GR"]
  meijing_hg19_GR[,cond := c("IMR90","K562","U20S")]
  carroll_hg19_ER[,Organism := "Human"]  
  carroll_hg19_ER[,Replicate := c(1,3,2)]
  carroll_hg19_ER[,TF := "ER"]
  carroll_hg19_ER[ ,cond := "MCF-7"]
  venters_hg19_TBP[,Organism := "Human"]
  venters_hg19_TBP[,cond := "K562"]
  venters_hg19_TBP[,Replicate := 1:3]
  venters_hg19_TBP[,TF := "TBP"]

  ## this creates the table
  A <- do.call(rbind,list(landick_aero,
                     landick_rif[order(Replicate)],
                     pugh_ctcf[order(Replicate)],
                     carroll_mm9_FoxA1[order(Replicate)],
                     meijing_hg19_GR[order(Replicate)],
                     carroll_hg19_ER[order(Replicate)],
                     venters_hg19_TBP[order(Replicate)]))
  A[,nsc := unlist(nsc)]
  setcolorder(A, c("Organism","TF","cond","Replicate","nreads","pbc","nsc","files"))
  A[,files := NULL]
  A[,nreads := prettyNum(nreads,big.mark = ",")]
  A[,pbc := round(pbc,4)]
  A[,nsc := round(nsc,4)]
  setnames(A,names(A),c("Organism","IP/TF","Condition/Cell","Rep.","Depth","PBC","NSC"))
@

@@

\subsubsection{Analysis of ChIP-Nexus data with the QC pipeline}

ChIP-nexus developed by He et al., 2015 \cite{chipnexus} is a
modification of the ChIP-exo protocol, where both sequencing adaptors
are ligated at the $3\prime$ end of the ChIP fragments. Then,
exonuclease digestion and DNA self-circularization with circLigase,
and restriction enzyme cutting between the two adaptors to create the
final library that is amplified.

@@

@@

For every ChIP-exo experiment, we partitioned the experiment into the
non-digested regions using the QC pipeline. For each region we
calculated the FSR defined as the ratio between the number of forward
stranded reads and the total number of reads in a region.

Then we
transformed the FSR into an imbalance index is defined as:

@@

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width = .46\textwidth]{../figs/for_paper/sensitivity_exo_olda_data.pdf}
  \includegraphics[width = .46\textwidth]{../figs/for_paper/resolution_by_dataset_old_data.pdf}
   \caption{Comparison of (A) sensitivity and (B) resolution between
     ChIP-exo and ChIP-Seq data. Sensitivity is defined as the
     proportion of RegulonDB annotations identified using each
     data. Resolution is defined as the distance between RegulonDB
     annotation and its closest prediction.}
  \label{fig:reso_all}
\end{figure}

@@

\begin{table}[h!]
%% need to fix format
  \centering
<<table_chipseq1,include = TRUE,echo = FALSE,eval = TRUE>>=
  load("../data/paper_tables/Landick_rif_chipseq_summary.RData")
  do.call(rbind,chipseq)
@ 
\caption{Same QC metrics as in table \ref{tab:qc} but applied to
  Landick's chipseq data of the rif experiment}
\end{table}

@@

>>>>

    ChIP-exo is a modification of the ChIP-Seq protocol for high
    resolution mapping of transcription factor binding sites. Although
    many aspect of the ChIP-exo data analysis are similar to those of
    ChIP-Seq, ChIP-exo presents a number of unique challenges. We
    present a quality control pipeline that analyzes a ChIP-exo and
    ChIP-nexus experiment's strand imbalance, enrichment and library
    complexity. Assessment of these characteristics are facilitated
    through diagnostic plots and summary statistics calculated over
    regions of the genome with varying levels of coverage.

@@

\begin{keyword}
  \kwd{ChIP-exo}
  \kwd{ChIP-nexus}
  \kwd{ChIP-Seq}
  \kwd{Quality Control}
  \kwd{Spatial Resolution}
  \kwd{Transcription Factor}  
  \kwd{Binding Site Identification on High-Res}
  \kwd{Deconvolution}
\end{keyword}

@@

% \newpage

% \tableofcontents
% \listoffigures

@@

% Refers to... First, DNA libraries... I think that these might be
% too much details for the introduction. How about discussing them in
% the results when we investigate the related

@@

\begin{table}[h!]
%% need to fix format
  \centering
<<table1,include = FALSE,echo = FALSE,eval = TRUE>>=
  library(data.table)
  load("../data/paper_tables/Landick_rif_summary.RData")
  load("../data/paper_tables/Landick_aero_summary.RData")
  load("../data/paper_tables/pugh_ctcf.RData")
  load("../data/paper_tables/Carroll_mm9_FoxA1.RData")
  load("../data/paper_tables/Meijsing_hg19_GR.RData")
  load("../data/paper_tables/Carroll_hg19_ER.RData")
  load("../data/paper_tables/Venters_hg19_TBP.RData")
  load("../data/paper_tables/ChIPnexus_dm3.RData")
  load("../data/paper_tables/ChIPnexus_hg19.RData")
  landick_aero[,Organism := "E.Coli"]
  landick_aero[,Replicate := c(1,2,1,2)]
  landick_aero[,TF := "Sig70"]
  landick_aero[ ,cond := c(rep("Aerobic",2),rep("Anaerobic", 2))]
  landick_rif[,Organism := "E.Coli"]
  landick_rif[,Replicate := c(1,1,2,2)]
  landick_rif[,TF := "Sig70"]
  landick_rif[ ,cond := rep(c("Rif-0min","Rif-20min"),each = 2)]
  pugh_ctcf[,Organism := "Human"]
  pugh_ctcf[,Replicate := 1]
  pugh_ctcf[,TF := "CTCF"]
  pugh_ctcf[ ,cond := "HeLA"]
  carroll_mm9_FoxA1[,Organism := "Mouse"]
  carroll_mm9_FoxA1[,Replicate := c(3,1,2)]
  carroll_mm9_FoxA1[,TF := "Fox A1"]
  carroll_mm9_FoxA1[ ,cond := "Mouse Liver"]
  meijing_hg19_GR[,Organism := "Human"]
  meijing_hg19_GR[,Replicate := 1]
  meijing_hg19_GR[, TF := "GR"]
  meijing_hg19_GR[,cond := c("IMR90","K562","U20S")]
  carroll_hg19_ER[,Organism := "Human"]  
  carroll_hg19_ER[,Replicate := c(1,3,2)]
  carroll_hg19_ER[,TF := "ER"]
  carroll_hg19_ER[ ,cond := "MCF-7"]
  venters_hg19_TBP[,Organism := "Human"]
  venters_hg19_TBP[,cond := "K562"]
  venters_hg19_TBP[,Replicate := 1:3]
  venters_hg19_TBP[,TF := "TBP"]
  chipnexus_dm3[, Organism := "D. Melanogaster"]
  chipnexus_dm3[, cond := c(rep("Embryo",4),rep("S2",4))]
  chipnexus_dm3[, Replicate := rep(1:2,4)]
  chipnexus_dm3[, TF := c(rep("Dorsal",2),rep("Twist",2),rep("Max",2),rep("MyC",2))]
  chipnexus_K562[ , Organism := "Human"]
  chipnexus_K562[ , cond  := "K562"]
  chipnexus_K562[ , Replicate := 1:2]
  chipnexus_K562[ , TF := "TBP"]
  ## this creates the table
  A <- do.call(rbind,list(landick_aero,
                     landick_rif[order(Replicate)],
                     pugh_ctcf[order(Replicate)],
                     carroll_mm9_FoxA1[order(Replicate)],
                     meijing_hg19_GR[order(Replicate)],
                     carroll_hg19_ER[order(Replicate)],
                     venters_hg19_TBP[order(Replicate)]))
  B <- do.call(rbind,list(chipnexus_dm3,
                     chipnexus_K562))
  format_tab <- function(A){
    A[,nsc := unlist(nsc)]
    setcolorder(A, c("Organism","TF","cond","Replicate","nreads","pbc","nsc","files"))
    A[,files := NULL]
    A[,nreads := prettyNum(nreads,big.mark = ",")]
    A[,pbc := round(pbc,4)]
    A[,nsc := round(nsc,4)]
    setnames(A,names(A),c("Organism","IP/TF","Condition/Cell","Rep.","Depth","PBC","NSC"))
    return(A)}

  A <- format_tab(A)
  B <- format_tab(B)
@

@@

\subsubsection{Analysis of ChIP-nexus data with the QC pipeline}

ChIP-nexus developed by He et al., 2015 \cite{chipnexus} is a
modification to the ChIP-exo protocol, where both sequencing adaptors
are ligated at the $3\prime$ end of the ChIP fragments. Then, after
exonuclease digestion, DNA self-circularization with circLigase, and
restriction enzyme cutting between the two adaptors, the final library
is amplified. In Table \ref{tab:nexus} we show the usual ChIP-Seq QC
metrics to ChIP-nexus data and found that the conclusions made for
ChIP-exo data still holds, the NSC is high but when plotting the SCC
the fragment and read length peaks are confounded and the PBC may
indicate severe bottlenecking but it is caused by the exonuclease
enzyme digestion which reduces the number of positions where the
reads can be aligned.

\begin{table}[h!]
  \centering
<<table3,include = TRUE,echo = FALSE, eval = TRUE >>= 
  B
@ 
\caption{\textbf{Current QC metrics applied to ChIP-nexus data.} PBC
  stands for PCR Bottleneck Coefficient and NSC for Normalized Strand
  Cross-Correlation.}
  \label{tab:nexus}
\end{table}

We applied the QC pipeline to the ChIP-nexus experiments and found
that the overall pattern seen in the visualizations is similar to a
high quality ChIP-exo experiment. Figure \ref{fig:nexus}A illustrates
a typical pattern of the ARC vs. URCR plot with ChIP-nexus data: Both
arms seems to be more narrow and therefore the separation between them
seems to be more noticeable, this seems reasonable since libraries
generated by ChIP-nexus usually yields higher complexity than
libraries generated by ChIP-exo. Additionally, while in ChIP-exo
experiments is common to observe regions formed by undigested
fragments in only one strand, we can observe that in ChIP-nexus
experiment the occurrence of this type of regions is reduced. Figure
\ref{fig:nexus}B shows that in a ChIP-nexus experiment the FSR
distribution approximates the median more quickly than in a ChIP-exo
experiment and Figure \ref{fig:nexus}C we observe that most of these
single-stranded regions are formed by few fragments, which means that
good ChIP-nexus experiment are less likely to show very imbalanced
regions.

\subsection{Evaluating ChIP-exo and ChIP-nexus data}

<<all_param,include = FALSE,echo = FALSE,eval = TRUE>>=

  Ntimes <- prettyNum(1e4,big.mark = ",")
  Msamp <- prettyNum(1e3,big.mark = ",")
  thresh1 <- 10

We applied the QC pipeline to all ChIP-exo and ChIP-nexus datasets
listed in Tables \ref{tab:qc} and \ref{tab:nexus}, that way we
partitioned the reads of each experiment into the undigested regions
by keeping the regions formed by at least one fragment. Figure
\ref{fig:enrich_all}A shows compares the slope of the following model:

\begin{align}
  \mbox{depth} = \beta \times \text{(\# of unique positions)} +\epsilon  \nonumber
\end{align}

when sampling \Sexpr{Msamp} regions and repeating this process
\Sexpr{Ntimes} times for every ChIP-exo and ChIP-nexus experiment. The
slope can be interpreted as the change in depth as number of unique
position varies. This parameter measures the library complexity on a
ChIP-exo experiment, since when it is high it means that several
fragments are being aligned to the same position in the genome.

Figures \ref{fig:enrich_all}B and \ref{fig:enrich_all}C shows the
parameters of the following model:

\begin{align}
  \mbox{URCR} = \frac{\kappa}{\mbox{ARC}} + \eta + \epsilon \nonumber
\end{align}

considering only the regions composed by 10 or more aligned reads,
that way the model reflects the enriched arm of the ARC vs. URCR
visualization. The intercept $\eta$ can be interpreted as the
theoretical $\mbox{URCR}$ for relatively high depth regions and the
slope $\kappa$ can be interpreted as the decay in $\mbox{URCR}$ as the
$\mbox{ARC}$ increases. That way experiments with low complexity are
showing low values for both parameters. We considered a linear model
between the estimated parameters for each experiment and it's
respective depth for which we were able to asses that there is not
significant relationship between the estimated parameters and the
interaction between depth and organism for each experiment. Hence,
these parameters may be used as indicators of the overall quality of
ChIP-exo experiment, but as with the SCC it is recommendable to also
visualize the SCC curve.

@@

\color{red}
\emph{need to add this part}
\color{black}

@@

For every ChIP-exo experiment, we partitioned the experiment into the
non-digested regions using the QC pipeline. For each region we
calculated the FSR defined as the ratio between the number of forward
stranded reads and the total number of reads in a region. Then we
transformed the FSR into an imbalance index is defined as:

@@

We used the imbalance index to asses if the regions considered by the
QC pipeline are less imbalanced when they overlap peak regions. Since
the strand imbalance distribution for both classes have the same
median and differ in the tails, this transformation separates the
tails from the median. Hence, when we reject the hypothesis of the
distribution being the same for both classes, it means that the class
of region that don't overlap peaks shows heavier tails than its
complement. \footnote{not sure if I should add the FSR densities for
  both classes in the supplement to prove this statement}

\subsection*{Evaluation of QC visualizations}

{\color{blue} \emph{need to describe figure \ref{fig:enrich_all}
}
\color{black}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width = .8\textwidth]{../figs/ChIPnexus/ChIPnexus_S2_MyC_enrichment.pdf}
\newline
  \includegraphics[width = .4\textwidth]{../figs/ChIPnexus/ChIPnexus_S2_MyC_fsr_surv.pdf}  
  \includegraphics[width = .4\textwidth]{../figs/ChIPnexus/ChIPnexus_S2_MyC_label_surv.pdf}  
  \caption{\textbf{ChIP-exo QC pipeline on ChIP-nexus data.} A)
    Enrichment vs. library complexity, B) Strand imbalance and C)
    Single-strand region analysis for the MyC experiment in S2 cell
    lines in \emph{D. Melanogaster} from He et al., 2015
    \cite{chipnexus}}
  \label{fig:nexus}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width = .9  \textwidth,page = 1]{../figs/test_quantify/slope_npos_depth_boxplot.pdf}
  \includegraphics[width = .9  \textwidth,page = 1]{../figs/test_quantify/arc_urcr_param_regression.pdf}
  \includegraphics[width = .9  \textwidth,page = 2]{../figs/test_quantify/arc_urcr_param_regression.pdf}
  \caption{A) Comparison of change in depth as the number of unique
    position obtained by sampling \Sexpr{Msamp} regions and fitting
    the model $\mbox{depth} = \beta \times \text{(\# unique
      positions)}$, and repeating this process \Sexpr{Ntimes} times. B)
    Intercept and C) Slope in the the $\mbox{URCR} = k / \mbox{ARC} +
    b$ model for the regions formed by more than 10 aligned
    positions.}
  \label{fig:enrich_all}
\end{figure}

\newpage
\begin{figure}[h!]
  \centering
  \includegraphics[width = .46\textwidth]{../figs/for_paper/sensitivity_exo_olda_data.pdf}
  \includegraphics[width = .46\textwidth]{../figs/for_paper/resolution_by_dataset_old_data.pdf}
   \caption{Comparison of (A) sensitivity and (B) resolution between
     ChIP-exo and ChIP-Seq data. Sensitivity is defined as the
     proportion of RegulonDB annotations identified using each
     data. Resolution is defined as the distance between RegulonDB
     annotation and its closest prediction.}
  \label{fig:reso_all}
\end{figure}

@@

\begin{table}[h!]
%% need to fix format
  \centering
<<table_chipseq1,include = TRUE,echo = FALSE,eval = TRUE>>=
  load("../data/paper_tables/Landick_rif_chipseq_summary.RData")
  chipseq$set <- chipseq$set[,Organism := "E.Coli"]
  chipseq$set <- chipseq$set[,Replicate := c(1,1,2,2)]
  chipseq$set <- chipseq$set[,TF := "Sig70"]
  chipseq$set <- chipseq$set[ ,cond := rep(c("Rif-0min","Rif-20min"),each = 2)]
  chipseq$pet <- chipseq$pet[,Organism := "E.Coli"]
  chipseq$pet <- chipseq$pet[,Replicate := c(1,1,2,2)]
  chipseq$pet <- chipseq$pet[,TF := "Sig70"]
  chipseq$pet <- chipseq$pet[ ,cond := rep(c("Rif-0min","Rif-20min"),each = 2)]
  chipseq <- lapply(chipseq,format_tab)
  chipseq <- do.call(rbind,chipseq)
  chipseq <- chipseq[, Protocol := rep(c("PE","SE"),each = 4)]
  chipseq
@ 
\caption{Same QC metrics as in table \ref{tab:qc} but applied to
  Landick's chipseq data of the rif experiment}
\end{table}

@@

% LocalWords:  Ntimes

### 2016-04-12 rmarkdown/paper.tex
preliminary version of the paper, pre -change of sections

\subsubsection{Analysis of ChIP-Nexus data with the QC pipeline}

ChIP-nexus developed by He et al., 2015 \cite{chipnexus} is a
modification of the ChIP-exo protocol, where both sequencing adaptors
are ligated at the $3\prime$ end of the ChIP fragments. Then,
exonuclease digestion and DNA self-circularization with circLigase,
and restriction enzyme cutting between the two adaptors to create the
final library that is amplified.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width = .46\textwidth]{../figs/for_paper/sensitivity_exo_olda_data.pdf}
  \includegraphics[width = .46\textwidth]{../figs/for_paper/resolution_by_dataset_old_data.pdf}
   \caption{Comparison of (A) sensitivity and (B) resolution between
     ChIP-exo and ChIP-Seq data. Sensitivity is defined as the
     proportion of RegulonDB annotations identified using each
     data. Resolution is defined as the distance between RegulonDB
     annotation and its closest prediction.}
  \label{fig:reso_all}
\end{figure}

@@

\begin{table}[h!]
%% need to fix format
  \centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{verbatim}
##                      files   nreads       pbc      nsc
## 1: edsn1396_Sig70.sort.bam 13445022 0.9426179 8.865244
## 2: edsn1398_Sig70.sort.bam 16538920 0.9378843 7.031836
## 3: edsn1400_Sig70.sort.bam 11642722 0.8891744 10.77284
## 4: edsn1402_Sig70.sort.bam 16854026 0.9407020 7.936239
## 5: edsn1396_Sig70.sort.bam  6722511 0.6632742  9.01779
## 6: edsn1398_Sig70.sort.bam  8269460 0.5594449 7.179539
## 7: edsn1400_Sig70.sort.bam  5821361 0.6472382 10.89898
## 8: edsn1402_Sig70.sort.bam  8427013 0.5895118 8.124717
\end{verbatim}
\end{kframe}
\end{knitrout}
\caption{Same QC metrics as in table \ref{tab:qc} but applied to
  Landick's chipseq data of the rif experiment}
\end{table}

>>>>

\subsubsection{Analysis of ChIP-nexus data with the QC pipeline}

ChIP-nexus developed by He et al., 2015 \cite{chipnexus} is a
modification to the ChIP-exo protocol, where both sequencing adaptors
are ligated at the $3\prime$ end of the ChIP fragments. Then, after
exonuclease digestion, DNA self-circularization with circLigase, and
restriction enzyme cutting between the two adaptors, the final library
is amplified. In Table \ref{tab:nexus} we show the usual ChIP-Seq QC
metrics to ChIP-nexus data and found that the conclusions made for
ChIP-exo data still holds, the NSC is high but when plotting the SCC
the fragment and read length peaks are confounded and the PBC may
indicate severe bottlenecking but it is caused by the exonuclease
enzyme digestion which reduces the number of positions where the
reads can be aligned.

\begin{table}[h!]
  \centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{verbatim}
##            Organism  IP/TF Condition/Cell Rep.       Depth    PBC     NSC
##  1: D. Melanogaster Dorsal         Embryo    1   8,863,170 0.6766  7.2771
##  2: D. Melanogaster Dorsal         Embryo    2  10,003,562 0.5657  7.1958
##  3: D. Melanogaster  Twist         Embryo    1  18,244,203 0.6592  5.8241
##  4: D. Melanogaster  Twist         Embryo    2  52,546,982 0.4550  5.2775
##  5: D. Melanogaster    Max             S2    1  18,320,743 0.5178  3.6057
##  6: D. Melanogaster    Max             S2    2  24,965,642 0.2124  3.4701
##  7: D. Melanogaster    MyC             S2    1   7,832,034 0.3935  5.9203
##  8: D. Melanogaster    MyC             S2    2  22,824,467 0.1879  5.7645
##  9:           Human    TBP           K562    1  33,708,245 0.3102 32.1617
## 10:           Human    TBP           K562    2 129,675,001 0.0492 32.7037
\end{verbatim}
\end{kframe}
\end{knitrout}
\caption{\textbf{Current QC metrics applied to ChIP-nexus data.} PBC
  stands for PCR Bottleneck Coefficient and NSC for Normalized Strand
  Cross-Correlation.}
  \label{tab:nexus}
\end{table}

We applied the QC pipeline to the ChIP-nexus experiments and found
that the overall pattern seen in the visualizations is similar to a
high quality ChIP-exo experiment. Figure \ref{fig:nexus}A illustrates
a typical pattern of the ARC vs. URCR plot with ChIP-nexus data: Both
arms seems to be more narrow and therefore the separation between them
seems to be more noticeable, this seems reasonable since libraries
generated by ChIP-nexus usually yields higher complexity than
libraries generated by ChIP-exo. Additionally, while in ChIP-exo
experiments is common to observe regions formed by undigested
fragments in only one strand, we can observe that in ChIP-nexus
experiment the occurrence of this type of regions is reduced. Figure
\ref{fig:nexus}B shows that in a ChIP-nexus experiment the FSR
distribution approximates the median more quickly than in a ChIP-exo
experiment and Figure \ref{fig:nexus}C we observe that most of these
single-stranded regions are formed by few fragments, which means that
good ChIP-nexus experiment are less likely to show very imbalanced
regions.

\subsection{Evaluating ChIP-exo and ChIP-nexus data}

We applied the QC pipeline to all ChIP-exo and ChIP-nexus datasets
listed in Tables \ref{tab:qc} and \ref{tab:nexus}, that way we
partitioned the reads of each experiment into the undigested regions
by keeping the regions formed by at least one fragment. Figure
\ref{fig:enrich_all}A shows compares the slope of the following model:

\begin{align}
  \mbox{depth} = \beta \times \text{(\# of unique positions)} +\epsilon  \nonumber
\end{align}

when sampling 1,000 regions and repeating this process
10,000 times for every ChIP-exo and ChIP-nexus experiment. The
slope can be interpreted as the change in depth as number of unique
position varies. This parameter measures the library complexity on a
ChIP-exo experiment, since when it is high it means that several
fragments are being aligned to the same position in the genome.

Figures \ref{fig:enrich_all}B and \ref{fig:enrich_all}C shows the
parameters of the following model:

\begin{align}
  \mbox{URCR} = \frac{\kappa}{\mbox{ARC}} + \eta + \epsilon \nonumber
\end{align}

considering only the regions composed by 10 or more aligned reads,
that way the model reflects the enriched arm of the ARC vs. URCR
visualization. The intercept $\eta$ can be interpreted as the
theoretical $\mbox{URCR}$ for relatively high depth regions and the
slope $\kappa$ can be interpreted as the decay in $\mbox{URCR}$ as the
$\mbox{ARC}$ increases. That way experiments with low complexity are
showing low values for both parameters. We considered a linear model
between the estimated parameters for each experiment and it's
respective depth for which we were able to asses that there is not
significant relationship between the estimated parameters and the
interaction between depth and organism for each experiment. Hence,
these parameters may be used as indicators of the overall quality of
ChIP-exo experiment, but as with the SCC it is recommendable to also
visualize the SCC curve.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width = .8\textwidth]{../figs/ChIPnexus/ChIPnexus_S2_MyC_enrichment.pdf}
\newline
  \includegraphics[width = .4\textwidth]{../figs/ChIPnexus/ChIPnexus_S2_MyC_fsr_surv.pdf}  
  \includegraphics[width = .4\textwidth]{../figs/ChIPnexus/ChIPnexus_S2_MyC_label_surv.pdf}  
  \caption{\textbf{ChIP-exo QC pipeline on ChIP-nexus data.} A)
    Enrichment vs. library complexity, B) Strand imbalance and C)
    Single-strand region analysis for the MyC experiment in S2 cell
    lines in \emph{D. Melanogaster} from He et al., 2015
    \cite{chipnexus}}
  \label{fig:nexus}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width = .9  \textwidth,page = 1]{../figs/test_quantify/slope_npos_depth_boxplot.pdf}
  \includegraphics[width = .9  \textwidth,page = 1]{../figs/test_quantify/arc_urcr_param_regression.pdf}
  \includegraphics[width = .9  \textwidth,page = 2]{../figs/test_quantify/arc_urcr_param_regression.pdf}
  \caption{A) Comparison of change in depth as the number of unique
    position obtained by sampling 1,000 regions and fitting
    the model $\mbox{depth} = \beta \times \text{(\# unique
      positions)}$, and repeating this process 10,000 times. B)
    Intercept and C) Slope in the the $\mbox{URCR} = k / \mbox{ARC} +
    b$ model for the regions formed by more than 10 aligned
    positions.}
  \label{fig:enrich_all}
\end{figure}

\newpage
\begin{figure}[h!]
  \centering
  \includegraphics[width = .46\textwidth]{../figs/for_paper/sensitivity_exo_olda_data.pdf}
  \includegraphics[width = .46\textwidth]{../figs/for_paper/resolution_by_dataset_old_data.pdf}
   \caption{Comparison of (A) sensitivity and (B) resolution between
     ChIP-exo and ChIP-Seq data. Sensitivity is defined as the
     proportion of RegulonDB annotations identified using each
     data. Resolution is defined as the distance between RegulonDB
     annotation and its closest prediction.}
  \label{fig:reso_all}
\end{figure}

@@

\begin{table}[h!]
%% need to fix format
  \centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{verbatim}
##    Organism IP/TF Condition/Cell Rep.      Depth    PBC     NSC Protocol
## 1:   E.Coli Sig70       Rif-0min    1 13,445,022 0.9426  8.8652       PE
## 2:   E.Coli Sig70       Rif-0min    1 16,538,920 0.9379  7.0318       PE
## 3:   E.Coli Sig70      Rif-20min    2 11,642,722 0.8892 10.7728       PE
## 4:   E.Coli Sig70      Rif-20min    2 16,854,026 0.9407  7.9362       PE
## 5:   E.Coli Sig70       Rif-0min    1  6,722,511 0.6633  9.0178       SE
## 6:   E.Coli Sig70       Rif-0min    1  8,269,460 0.5594  7.1795       SE
## 7:   E.Coli Sig70      Rif-20min    2  5,821,361 0.6472 10.8990       SE
## 8:   E.Coli Sig70      Rif-20min    2  8,427,013 0.5895  8.1247       SE
\end{verbatim}
\end{kframe}
\end{knitrout}
\caption{Same QC metrics as in table \ref{tab:qc} but applied to
  Landick's chipseq data of the rif experiment}
\end{table}
